---
title: "Project 1"
author: "Sunanda Das"
date: '2022-04-07'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing packages
```{r}
library(tidyverse) # metapackage of all tidyverse packages
library(ggcorrplot)
library(rsample)
```

#Reading in files
```{r}
df = read_csv("C:/Users/sdas/Desktop/Spring 2022/STA 6933/Project 1/diabetes.csv" , show_col_types = FALSE)
```

# Inspect the data
```{r}
colnames(df)[9] <- "Diabetes" #rename output as diabetes
colnames(df)[7] <- "Ped"  #rename DiabetesPedigreeFunction as Ped
colnames(df)[3] <- "BP"  #rename BloodPressure as BP
head(df,5)
```

```{r}
r <- cor(df, use="complete.obs")
round(r,2)
ggcorrplot(r, 
           hc.order = TRUE, 
           type = "lower",
           lab = TRUE)
```
Seems as if the highest correlation to our response variable is glucose with a correlation of 0.47. No strong correlation observed between variables. So, no need to drop any of them for analysis.


#Missing values
```{r}
# Treat 0 in the biological variables other than number of times pregnant as missing values 
cols_change <- colnames(df)[!colnames(df) %in% c("Pregnancies","Age","Ped", "Diabetes")]
data <- df[cols_change] == 0
df[cols_change][data] <- NA

# Show the number of missing values of each column
print(apply(data, 2, sum))
```

#Data Split

```{r}
# Set a random seed
set.seed(123)
# Split the dataset: 75% for trainging and 25% for testing

traindata <- sample(nrow(df), round(0.75*nrow(df)), replace = FALSE)
train <- df[traindata,]
test <- df[-(traindata),]
train$Diabetes<-as.numeric(train$Diabetes)
test$Diabetes<-as.numeric(test$Diabetes)
```


```{r}
prop.table(table(train$Diabetes))
```
```{r}
prop.table(table(test$Diabetes))
```
#Missing Data Imputation

Because of the small size of the dataset, I want to obtain as much as information from it, so I will not delete either entire observations (rows) or variables (columns) containing missing values right now.
Two options:
    Replacing missing data with sensible values (mean or median) given the distribution of the data.
    Replacing missing data with prediction (Mutiple Imputaion).

I’ll use 1st (median) on the small number of missing values and 2nd (Mutiple Imputaion) on the large number of missing values.

```{r}
# Median value imputation
train$Glucose[is.na(train$Glucose)] <- median(train$Glucose,na.rm = T)
train$BP[is.na(train$BP)] <- median(train$BP,na.rm = T)
train$BMI[is.na(train$BMI)] <- median(train$BMI,na.rm = T)

# Multiple imputation
library(mice)
mice_mod <- mice(train[, c("SkinThickness","Insulin")], method='rf') 
```


```{r}
# Save the complete imputation output 
mice_complete <- complete(mice_mod)

# Show distributions for skinThickness and Insulin
par(mfrow=c(2,2))
hist(train$SkinThickness, freq=F, main='Triceps skin fold thickness : Original Data',
     col='darkgreen', ylim=c(0,0.04))
hist(mice_complete$SkinThickness, freq=F, main='Triceps skin fold thickness : MICE Output',
     col='lightgreen', ylim=c(0,0.04))
hist(train$Insulin, freq=F, main='2-Hour serum insulin: Original Data',
     col='darkblue', ylim=c(0,0.004))
hist(mice_complete$Insulin, freq=F, main='2-Hour serum insulin: MICE Output',
     col='lightblue', ylim=c(0,0.004))
```
Compared with the original distributions, the two complete distributions above for tsk thickness and serum are not significant changed, which is good.

```{r}
# Replace skinThickness and Insulin variables from the mice
train$SkinThickness <- mice_complete$SkinThickness
train$Insulin <- mice_complete$Insulin

# Make sure there is no missing data
sum(is.na(train))
```
#build models

#Data Normalization

One of the most important procedures when forming a neural network is data normalization. This involves adjusting the data to a common scale so as to accurately compare predicted and actual values. Failure to normalize the data will typically result in the prediction value remaining the same across all observations, regardless of the input values.

We can do this in two ways in R:

    Scale the data frame automatically using the scale function in R
    Transform the data using a max-min normalization technique

    
```{r}
# Normalize training data, Scaled Normalization
scale_training <- as.data.frame(scale(train[,-9],  
                                      center = TRUE, scale = TRUE))
scale_training$Diabetes<-as.numeric(train$Diabetes)
str(scale_training)
```
#Neural Network

```{r}
library("neuralnet")
scale_training$Diabetes <- as.factor(ifelse(scale_training$Diabetes==0, "non-diabetic", "Diabetic")) #recode Diabetes
NN = neuralnet(Diabetes ~ ., scale_training, hidden = 2 , linear.output = TRUE, act.fct="logistic", rep=1)
plot(NN)
```

```{r}
library(ggplot2)
library(cvms)
library(tibble)
predict <- neuralnet::compute(NN, scale_training[,1:8])
Prediction <- round(predict$net.result[,1])
Actual <- train$Diabetes
cm<-table(Actual,Prediction)
cfm <- as_tibble(cm)
plot_confusion_matrix(cfm, 
                      target_col = "Actual", 
                      prediction_col = "Prediction",
                      counts_col = "n")
```
#testing accuracy of model
```{r}
library(caret)
confusionMatrix(cm)
```
```{r}
library(pROC)
rocTest_nn <- roc(train$Diabetes, as.numeric(Prediction))
plot(rocTest_nn, print.auc = T)
```

#TEST DATASET:

```{r}
# Normalize testing data, Scaled Normalization
scale_test <- as.data.frame(scale(test[,-9],  
                                      center = TRUE, scale = TRUE))
scale_test$Diabetes<-as.numeric(test$Diabetes)
str(scale_test)
```

```{r}
predicttest <- neuralnet::compute(NN, scale_test[,1:8])
Prediction_test <- round(predicttest$net.result[,1])
Target <- test$Diabetes
cm2<-table(Target,Prediction_test)
cfm2 <- as_tibble(cm2)
plot_confusion_matrix(cfm2, 
                      target_col = "Target", 
                      prediction_col = "Prediction_test",
                      counts_col = "n")
```
#testing accuracy of model
```{r}
confusionMatrix(cm2)
```
```{r}
rocTest<- roc(test$Diabetes, as.numeric(Prediction_test))
plot(rocTest, print.auc = T)
```



